### 2018-07-05 paper/3-abstractions.tex
Address suggestions of reviewers
Polishing (thanks, Andrew!)

\noindent
A \hs{Task} describes a single build task, while \hs{Tasks} associates a
\hs{Task} to every non-input key; input keys are associated with
\hs{Nothing}. The highly-abstracted type\footnote{Readers familiar with
\emph{lenses} or \emph{profunctor optics} might recognise a familiar pattern.
We discuss this in~\S\ref{sec-related-optics}.} \hs{Task} is best introduced by
an example. Consider this \Excel spreadsheet:

@@

% \newpage

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Applicative functors
pure  :: Applicative f => a -> f a
(<$>) :: Applicative f =>   (a -> b) -> f a -> f b -- Left-associative
(<*>) :: Applicative f => f (a -> b) -> f a -> f b -- Left-associative
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard State monad from Control.Monad.State
data State s a
instance Monad (State s)
get       :: State s s
gets      :: (s -> a) -> State s a
put       :: s -> State s ()
modify    :: (s -> s) -> State s ()
runState  :: State s a -> s -> (a, s)
execState :: State s a -> s -> s
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard types from Data.Functor.Identity and Data.Functor.Const
newtype Identity a = Identity { runIdentity :: a }
newtype Const m a  = Const    { getConst    :: m }
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
instance Functor (Const m) where
    fmap _ (Const m) = Const m
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
instance Monoid m => Applicative (Const m) where
    pure _              = Const mempty
    Const x <*> Const y = Const (x <> y)
\end{minted}
\vspace{-3mm}
\caption{Standard library definitions.}\label{fig-stdlib}
\vspace{-4mm}
\end{figure}

@@

\noindent
Notice that these calls to \hs{dependencies} do no actual computation (in this
case, spreadsheet arithmetic). They cannot: we are not supplying a store or any
input numbers. So, through the wonders of polymorphism, we are able to extract
the dependencies of the spreadsheet formula, and to do so efficiently, simply by
running its code in a different \hs{Applicative}! This is not new, for example
see~\cite{free-applicatives}, but it is extremely cool. We will see a practical
use for \hs{dependencies} when implementing applicative build systems,
see~\S\ref{sec-implementation-make}.

>>>>

\noindent
% Here \hs{c} stands for \emph{constraint}, such as \hs{Applicative} or \hs{Monad}.
A \hs{Task} describes a single build task, while \hs{Tasks} associates a
\hs{Task} to every non-input key; input keys are associated with
\hs{Nothing}. The highly-abstracted type\footnote{Readers
familiar with \emph{lenses} or \emph{profunctor optics} might recognise a
familiar pattern. We discuss this in~\S\ref{sec-related-optics}.} \hs{Task} is
best introduced by an example. Consider this \Excel spreadsheet:

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Applicative functors
pure  :: Applicative f => a -> f a
(<$>) :: Applicative f =>   (a -> b) -> f a -> f b -- Left-associative
(<*>) :: Applicative f => f (a -> b) -> f a -> f b -- Left-associative
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard State monad from Control.Monad.State
data State s a
instance Monad (State s)
get       :: State s s
gets      :: (s -> a) -> State s a
put       :: s -> State s ()
modify    :: (s -> s) -> State s ()
runState  :: State s a -> s -> (a, s)
execState :: State s a -> s -> s
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard types from Data.Functor.Identity and Data.Functor.Const
newtype Identity a = Identity { runIdentity :: a }
newtype Const m a  = Const    { getConst    :: m }
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
instance Functor (Const m) where
    fmap _ (Const m) = Const m
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
instance Monoid m => Applicative (Const m) where
    pure _              = Const mempty   -- mempty is the identity of the monoid m
    Const x <*> Const y = Const (x <> y) -- <> is the binary operation of the monoid m
\end{minted}
\vspace{-3mm}
\caption{Standard library definitions.}\label{fig-stdlib}
\vspace{-4mm}
\end{figure}

@@

@@

\noindent
Notice that these calls to \hs{dependencies} do no actual computation (in this
case, spreadsheet arithmetic). They cannot: we are not supplying a store or any
input numbers. So, through the wonders of polymorphism, we are able to extract
the dependencies of the spreadsheet formula, and to do so efficiently, simply by
running its code in a different \hs{Applicative}! This is not new, for example
see~\citet{free-applicatives}, but it is extremely cool. We will see a practical
use for \hs{dependencies} when implementing applicative build systems,
see~\S\ref{sec-implementation-make}.

### 2018-07-05 paper/7-related.tex
Polishing (thanks, Andrew!)

To achieve minimality, each separate command is traced at the OS-level, allowing
\Fabricate to record a trace entry stating that \cmd{gcc -c util.c} reads from
\cmd{util.c}. In future runs \Fabricate runs the script from start to finish,
skipping any commands where no inputs have changed. The key difference from our
\hs{Tasks} abstraction is that instead of supplying a mapping from output to action,
\Fabricate supplies a list of statements, in an order, without declaring what each line produces.
There is no need to schedule the statements, and not enough information to do so.

@@

%TODO: Add a reference to the paper about forall f. c f => (k -> f v) -> f a
\begin{minted}[xleftmargin=10pt]{haskell}
type Optic p a b s t = p a b -> p s t
\end{minted}
\noindent
provided we instantiate \hs{p}~\hs{a}~\hs{b} to something like
\hs{k}~\hs{->}~\hs{f}~\hs{v}~--~which many of the actual instances in that paper
do. The properties of such optics are well studied, and the functions like
\hs{dependencies} are very much based on observations from that field of work.
Alas, we have been unable to remove the \hs{Maybe} used to encode whether a file
is an input, without complicating other aspects of our definition. Furthermore,
the \hs{Build} abstraction lacks any further such symmetry.

>>>>

\noindent
To achieve minimality, each separate command is traced at the OS-level, allowing
\Fabricate to record a trace entry stating that \cmd{gcc -c util.c} reads from
\cmd{util.c}. In future runs \Fabricate runs the script from start to finish,
skipping any commands where no inputs have changed. The key difference from our
\hs{Tasks} abstraction is that instead of supplying a mapping from output to action,
\Fabricate supplies a list of statements, in an order, without declaring what each line produces.
There is no need to schedule the statements, and not enough information to do so.

@@

%TODO: Add a reference to the paper about forall f. c f => (k -> f v) -> f a
\begin{minted}[xleftmargin=10pt]{haskell}
type Optic p a b s t = p a b -> p s t
\end{minted}
\noindent
Provided we instantiate \hs{p}~\hs{a}~\hs{b} to something like
\hs{k}~\hs{->}~\hs{f}~\hs{v}~--~which many of the actual instances in that paper
do. The properties of such optics are well studied, and the functions like
\hs{dependencies} are very much based on observations from that field of work.
Alas, we have been unable to remove the \hs{Maybe} used to encode whether a file
is an input, without complicating other aspects of our definition. Furthermore,
the \hs{Build} abstraction lacks any further such symmetry.
